Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{ I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback. }

### 2023-06-26 paper/paper.tex
#29 finish adding Col (5) and feedback 3/3

As a first check I include different controls $\textbf{X}_{it}$ in Equation \eqref{eq:baseline}. Table \ref{tab:controls_50y} shows the results of this exercise. Column (1) is identical to Column (1) in Table \ref{tab:baseline_50y} and controls only for switching in the same period. In Column (2) I remove this control. Columns (3) and (4) add conflict incidents as a control variable, once in addition to switching and once without. The outcome is $All \, construction$ in all regressions.

@@

While this result may seem like evidence for omitted variable bias in the baseline estimate, that explanation is ultimately not plausible. To cause bias in Column (1) conflict incidents would need to be correlated both with the post-conquest dummy and with construction --- however, we can see that the latter is not the case: the coefficients on $Conflict$ in Columns (3) and (4) are tiny and not significant by any stretch of the imagination.\footnote
{
The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations. Simply computing the correlation coefficient between the dummies for conflict incidents and for construction yields a value of $-0.01$.
} 
Rather, I suggest that the estimates for $\theta_{conquest}$ are different because they are estimated on different samples. More than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable. Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space --- if the spatial composition of the smaller sample was the same, the entire map would be a uniform shade. Instead we see big discrepancies: the west, the south, and Brandenburg retain 30\% to 60\% of their observations, while many northern and eastern regions all but vanish from the sample.

If we accept that the disagreement between the interaction coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous differences in treatment effects. The null difference in (1) could be an average of the negative difference in the southwest and a hypothetical positive difference in the northeast. Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory: the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
{
Of course, England and Sweden would attack via the North and Baltic seas, but it seems a reasonable assumption that competition for territory would be more intense between states that share a land border.
} 
BLANK \cite{cervellati2022}. I leave this for future research to explore.

>>>>

As a first check I include different controls $\textbf{X}_{it}$ in Equation \eqref{eq:baseline}. Table \ref{tab:controls_50y} shows the results of this exercise. Column (1) is identical to Column (1) in Table \ref{tab:baseline_50y} and controls only for switching in the same period. In Column (2) I remove this control. Columns (3) and (4) add conflict incidents as a control variable, once in addition to switching and once without. Finally, Column (5) runs the baseline specification from (1) on the smaller sample used in (3) and (4). The outcome is $All \, construction$ in all regressions.

@@

While this result may seem like evidence for omitted variable bias in the baseline estimate, that explanation is ultimately not plausible. Column (5) shows that the disagreement between the estimates does not come from controlling for conflict, but from the sample composition: more than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable. When I run the baseline specification on this partial sample, I obtain the exact same estimates as in Column (3) --- even without controlling for conflict.

Figure \ref{fig:conflict_map} shows that the observations that remain in the partial sample follow quite a different distribution across space --- if the spatial composition of the partial sample was the same as that of the original sample, the entire map would be a uniform shade. Instead we see big discrepancies: the west, the south, and Brandenburg retain 30\% to 60\% of their observations, while many northern and eastern regions all but vanish from the sample.

Together, Table \ref{tab:controls_50y} and Figure \ref{fig:conflict_map} have the interesting implication of spatially heterogeneous differences in treatment effects. The null difference in Column (1) could be an average of the negative difference in the southwest and a hypothetical positive difference in the northeast. Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory: the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
{
Of course, England and Sweden would attack via the North and Baltic seas, but it seems a reasonable assumption that competition for territory would be more intense between states that share a land border.
} 
BLANK \cite{cervellati2022}. I leave this for future research to explore.
